Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-11-24 paper/main.tex
rw: bundle and sheaves

\usepackage[switch,columnwise]{lineno}

@@

\begin{definition}
The \textcolor{base}{base space} \dbasec\ is a topological space, which means it is a set $\dbase$ with a collection of open sets \cite{TopologyVsTopology} surrounding each element in the set. Open sets are a collection of subsets $\{\opensetc\}$ in a set \dbase, including the empty set, the whole set \dbase, and every union and intersection of its member subsets. For each point $\dbasepoint \in \dbase$, there is a collection of subsets $\\{\openset\} \subset \dbase$ such that $\dbasepoint \in \openset \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}  
\end{definition}
\note{intuition about base space}

@@

\begin{definition} The \textcolor{total}{total space} \dtotalc\ is the space reachable via the projection map $\pi$. The total space is always locally trivial, meaning $\dtotal = \dbase \times \dfiber$ over any open neighborhood $\openset_{\dbasepoint}$. 
\end{definition}
The fiber bundle is trivial when $\dtotal = \dbase \times \dfiber$ is true globally; a non-trivial bundle can be thought of as the twisted fiber product $\dtotal=\dbase times_{T} \dfiber$\cite{hatcherAlgebraicTopology2002,nlab:twisting_function}. The twisting refers to the fiber \dfiber\ not aligning in the same direction over all $\dopenspace \subseteq \dbase$.

\begin{equation}
  \label{eq:related-work:continuity:fiber-bundle}
  \begin{tikzcd}
      \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
                        & \dbasec \arrow[u, "\dsectionc"', bend right]
  \end{tikzcd}
\end{equation}

\note{this needs to be reworked into a coherent sentence}
How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$. In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....

As proposed by Butler et al, data can be represented as a map from the base space \dbase\ to the fiber space \dfiber. This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle. A fiber bundle has many sections, and the set of all sections
\begin{equation}
\label{eq:related-work:continuity:sections}
\{\dsection^{i}: \openset_{j}\rightarrow \dtotal\}|_{i\in I} = \Gamma(\openset_{j}, \dtotal)
\end{equation}
over a subset \openset\ is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase. The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.

\subsubsection{Sheaves}
The map from the open set to the set of sections is called a presheaf $\sheaf: U_j \rightarrow \Gamma(\openset_j, \dtotal)$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. The presheaf preserves inclusion maps between the open sets open sets $\iota$ and the inclusion map $\iota^*$ between the sets of sections.

\begin{equation}
  \label{eq:related-work:continuity:presheaf}
  \begin{tikzcd}[column sep=small]
    {\Gamma(\opensetc_1, \dtotalc)}                                       & {}                                                    & {\Gamma(\opensetc_2, \dtotalc)} \arrow[ll, "\iota^*"', hook'] \\
                                                                  &                                                       &                                                       \\
    \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    
    \end{tikzcd}
\end{equation}

This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subspace \openset_2$. Sheaves are presheafs where sets of sections $\Gamma$ are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dtotal$ \cite{spanier1989algebraic, nlab:sheaf}. Sheaves are often used as an abstraction for keeping track of how data over topological spaces is glued together \cite{ghristElementaryAppliedTopology2014} because they model the data as sets of sections, continuity of the data in the base, and how subsets fit together through inclusion maps.

@@

\begin{definition} 
  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
  \begin{itemize}
    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
  \end{itemize}
\end{definition}
The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{\fongInvitationAppliedCategory2019}. For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) | \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$. For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$. Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.

@@

The continuity of the graphic is encapsulated as a category $\mathcal{\mathcal{\gbase}}$ containing open set objects $\opensetg$, the complete set of opensets $\gbase$ that is the full space of the graphic, and the empty set. As with $\mathcal{\dbase}$, the morphisms between the open sets  $\opensetg$ are inclusion morphisms $\iota$. Over the graphic space $\opensetg \subseteq \gbase$ are a set of sections 
\begin{equation}
  \label{eq:artist:graphic_sections}
  \{\gsectionc^{i}:\opensetgc_i\rightarrow \gtotalc\}|_{i\in I}
\end{equation}
which are the set of all graphics with a shared base space and fiber and are a subcategory of \textbf{Set}. This allows us to represent the full space of graphics as a sheaf $\sheaf(\gtotal)$
\begin{equation}
  \label{eq:artist:artist_sheaf}
  \begin{tikzcd}
    {Ob(\bm{Set}) \ni \Gamma(\opensetgc_i, \gtotalc)}                                                                                                  &  & \{\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\}|_{j\in J} \arrow[ll, "\iota"', hook'] \\ &  & \\
    \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)"] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\gtotalc)" description, dashed] &  & 
    \end{tikzcd}
\end{equation}
The set of graphics reachable through an artist function is a subset $\{\gsection^{j}\}|_{j\in J}$ of the full set of graphics. These sets of reachable graphics over opensets \opensetg\ are also objects of \textbf{Set}. This allows us to formulate the graphic output by the artist as a sheaf $\sheaf_{\vartist}$ from the graphic base space to the set of reachable graphics.

To establish a mapping between graphic and data base spaces, we introduce the surjective map $\vindex: \gtotal \rightarrow \dtotal$. Given the morphism $\iota$
\begin{equation}
  \label{eq:artist:vindex}
  \begin{tikzcd}
    & \opensetgc_i \arrow[d, "\iota"] \\
  \dbasec \arrow[ru, "\vindexc^*", dashed] & \gbasec \arrow[l, "\vindexc"]   
\end{tikzcd}
\end{equation}
we propose that a pushforward functor $\vindex_*$ 
\begin{equation}
  \vindexc_{*}:\mathcal{\dbasec} \rightarrow \mathcal{\gbasec}
\end{equation}
can be used to map between open sets $\openset_i \subset \dbase$ in data space and the open sets in visual space $\opensetg_i \subset \gbase$. We can then compose the functor $\vindex_*$ with the sheaf $\sheaf(\dtotal)$ 
\begin{equation*}
\vindex_*(\openset) \rightarrow \Gamma(\vindex_*(\openset), \vindex_* \dtotal) \coloneqq \opensetg \rightarrow \gamma(\opensetg, \vindex_* \dtotal) 
\end{equation*}
which allows for moving the sets of sections from $\Gamma(\openset, \dtotal)$ to $\Gamma(\opensetg, \vindex_* \dtotal)$. We denote the pulled forward sheaf as $\sheaf_{\vindex_*}(\dtotal)$.

@@

of sheaf functors 
\begin{equation}
  \label{eq:artist:sheaf_map}
  \vartist: \sheafc_{vindex_*}(\dtotal) \rightarrow \sheafc_{\vartist}(\gtotal)
\end{equation}
from the graphic base space category $\mathcal{\gbase}^{op}$ to different objects in \bm{Set}. This means that the following diagram commutes

>>>>

\usepackage[]{lineno}

@@

\begin{definition}
The \textcolor{base}{base space} \dbasec\ is a topological space, which means it is a set $\dbase$ with a collection of open sets \cite{TopologyVsTopology} surrounding each element in the set. Open sets are a collection of subsets $\{\opensetc\}$ in a set \dbase, including the empty set, the whole set \dbase, and every union and intersection of its member subsets. For each point $\dbasepoint \in \dbase$, there is a collection of subsets $\{\openset\} \subset \dbase$ such that $\dbasepoint \in \openset \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}  
\end{definition}
\note{intuition about base space}

@@

\begin{definition} The \textcolor{total}{total space} \dtotalc\ is the space reachable via the projection map $\pi$. The total space is always locally trivial, meaning $\dtotal = \dbase \times \dfiber$ over any open neighborhood $\openset_{\dbasepoint}$. 
\end{definition}

The fiber bundle is trivial when $\dtotal = \dbase \times \dfiber$ is true globally; a non-trivial bundle can be thought of as the twisted fiber product $\dtotal=\dbase times_{T} \dfiber$\cite{hatcherAlgebraicTopology2002,nlab:twisting_function}. The twisting refers to the fiber \dfiber\ not aligning in the same direction over all $\openset \subseteq \dbase$.

As proposed by Butler et al \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}, data can be represented as a map from the base space \dbase\ to the fiber space \dfiber. This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle.

\begin{equation}
  \label{eq:related-work:continuity:fiber-bundle}
  \begin{tikzcd}[row sep=large]
      \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
                        & \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
  \end{tikzcd}
\end{equation}

A fiber bundle has many sections, and the set of all sections
\begin{equation}
\label{eq:related-work:continuity:sections}
\{\dsectionc^{i}: \opensetc\rightarrow \dtotalc\}|_{i\in I} = \cgamma{\opensetc}{\dtotalc}
\end{equation}
over an open set $openset \subseteq \dtotal$ is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase. The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.

\subsubsection{Sheaves}
Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces. A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \gammac{\openset}{ \dtotal}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. The presheaf preserves inclusion maps $\iota$ between open sets and the inclusion map $\iota^*$ between the sets of sections.

\begin{equation}
  \label{eq:related-work:continuity:presheaf}
  \begin{tikzcd}[column sep=small]
    {\cgamma{\opensetc_1}{\dtotalc}} & {} & {\cgamma{\opensetc_2}{\dtotalc}} \arrow[ll, "\iota^*"', hook'] \\
                                                                  &                                                       &                                                       \\
    \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    
    \end{tikzcd}
\end{equation}

This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$. Sheaves are presheafs where sets of sections $\Gamma$ are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dtotal$ \cite{spanier1989algebraic, nlab:sheaf}. Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.

@@

\begin{definition} 
  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
  \begin{itemize}
    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
  \end{itemize}
\end{definition}
The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}. For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) | \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$. For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$. Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.

@@

The continuity of the graphic is encapsulated as a category $\mathcal{\mathcal{\gbase}}$ containing open set objects $\opensetg$, the complete set of opensets $\gbase$ that is the full space of the graphic, and the empty set. As with $\mathcal{\dbase}$, the morphisms between the open sets  $\opensetg$ are inclusion morphisms $\iota$. Over the graphic space $\opensetg \subseteq \gbase$ are a set of sections 
\begin{equation}
  \label{eq:artist:graphic_sections}
  \{\gsectionc^{i}:\opensetgc_i\rightarrow \gtotalc\}|_{i\in I}
\end{equation}
which are the set of all graphics with a shared base space and fiber and are a subcategory of \textbf{Set}. This allows us to represent the full space of graphics as a sheaf $\sheaf(\gtotal)$
\begin{equation}
  \label{eq:artist:artist_sheaf}
  \begin{tikzcd}
    {Ob(\bm{Set}) \ni \Gamma(\opensetgc_i, \gtotalc)}                                                                                                  &  & \{\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\}|_{j\in J} \arrow[ll, "\iota"', hook'] \\ &  & \\
    \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)"] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\gtotalc)" description, dashed] &  & 
    \end{tikzcd}
\end{equation}
The set of graphics reachable through an artist function is a subset $\{\gsection^{j}\}|_{j\in J}$ of the full set of graphics. These sets of reachable graphics over open sets \opensetg\ are also objects of \textbf{Set}. This allows us to formulate the graphic output by the artist as a sheaf $\sheaf_{\vartist}$ from the graphic base space to the set of reachable graphics.

To establish a mapping between graphic and data base spaces, we introduce the surjective map $\vindex: \gtotal \rightarrow \dtotal$. Given the morphism $\iota$
\begin{equation}
  \label{eq:artist:vindex}
  \begin{tikzcd}
    & \opensetgc_i \arrow[d, "\iota"] \\
  \dbasec \arrow[ru, "\vindexc^*", dashed] & \gbasec \arrow[l, "\vindexc"]   
\end{tikzcd}
\end{equation}
we propose that a pushforward functor $\vindex_*$ 
\begin{equation}
  \vindexc_{*}:\mathcal{\dbasec} \rightarrow \mathcal{\gbasec}
\end{equation}
can be used to map between open sets $\openset_i \subset \dbase$ in data space and the open sets in visual space $\opensetg_i \subset \gbase$. We can then compose the functor $\vindex_*$ with the sheaf $\sheaf(\dtotal)$ 
\begin{equation*}
\vindex_*(\openset) \rightarrow \Gamma(\vindex_*(\openset), \vindex_* \dtotal) \coloneqq \opensetg \rightarrow \gamma(\opensetg, \vindex_*(\dtotal) 
\end{equation*}
which allows for moving the sets of sections from $\Gamma(\openset, \dtotal)$ to $\Gamma(\opensetg, \vindex_* \dtotal)$. We denote the pulled forward sheaf as $\sheaf_{\vindex_*}(\dtotal)$.

@@

of sheaf functors 
\begin{equation}
  \label{eq:artist:sheaf_map}
  \vartist: \sheafc_{vindex_*}(\dtotal) \rightarrow \sheafc_{\vartist}(\gtotal)
\end{equation}
where each sheaf is a map from an open set in the graphic base space $\opensetg \subset \gbase$ into a set of sections of a fiber bundle over that open set.

Because the artist is a natural transform, it also transforms morphisms between functors\cite{bradleyWhatNaturalTransformation, maclaneCategoriesWorkingMathematician2013}. As illustrated in \autoref{eq:artist:morphism-of-sheaf-data}, the morphism between data sheaves is $\dfunc$. By definition of a natural transform, the following diagram commutes
